### 2020-04-14 sections/discussion.tex
separates analysis of galaxy pitch angle and arm pitch angle

Our method and results do not completely reject spiral winding of the form described by \citet{2019arXiv190910291P}, however this result is highly sensitive to the boundaries used; using limits of $10.0^\circ < \phi < 40.0^\circ$ results in us rejecting the null hypothesis at the 1\% level for 99\% of the MCMC draws. As no physical justification was provided for the limits present in \citet{2019arXiv190910291P}, this is a far from ideal test of arm winding.

@@

>>>>

@@

Our results are consistent with spiral winding of the form described by \citet{2019arXiv190910291P}; providing evidence for transient and recurrent spiral arms, the evolution of which is governed by self-gravity (such as through swing-amplification, \citealt{1965MNRAS.130..125G})) and which wind up over time. The assumptions of this model of spiral winding are also highly reductionist, and leaves many unanswered questions: what determines the limits on $\phi$? Is the spiral arm equally apparent at all pitch angles, or is a selection effect present? This result is not evidence against QSDW, as it is possible that our distribtuion of pitch angles is dictated by other factors such as disk shear and central mass concentration.

### 2020-04-14 sections/introduction.tex
separates analysis of galaxy pitch angle and arm pitch angle

In this dynamic picture of spiral arms, pitch angle monotonically decreases from a spiral arm's formation to its dissapation. \citet{2019arXiv190910291P} proposes a simple test of spiral arm winding, assuming the cotangent of the pitch angle evolves linearly with time. Finding compelling evidence of this relationship against QSDW theory, in favour of the dynamic spirals now favoured by many simulations.

>>>>

In this dynamic picture of spiral arms, pitch angle monotonically decreases from a spiral arm's formation to its dissapation. \citet{2019arXiv190910291P} proposes a simple test of spiral arm winding, assuming the cotangent of the pitch angle of a spiral arm evolves linearly with time. They found that the pitch angles of their sample of 86 galaxies was consistent with this distribution, evidence against QSDW theory in favour of the dynamic spirals produced in many simulations.

### 2020-04-14 sections/method.tex
separates analysis of galaxy pitch angle and arm pitch angle

Spiral arm points are deprojected to a face-on orientation using the disk inclination and position angle obtained through photometric model fitting performed in \Lingard.

@@

We model a spiral arm as a logarithmic spiral, described by

\begin{equation}
r_\mathrm{arm} \sim \mathrm{Normal}\left(\exp\left[\theta_\mathrm{arm}\tan\phi_\mathrm{arm} + c_\mathrm{arm}\right], \sigma_r\right).
\end{equation}

Assume the pitch angle of a galaxy's spiral arms are drawn from a Normal distribution, truncated between 0 and 90 degrees, centred on some value we will call the galaxy's pitch angle, $\phi_\mathrm{gal}$. This dispersion in arm pitch angle has some measure of spread, $\sigma_\mathrm{gal}$, which we will assume is the same in all galaxies:

@@

We choose hyperpriors on $\phi_\mathrm{gal}$ and $\sigma_\mathrm{gal}$ of

\begin{align}
  \phi_\mathrm{gal} &\sim \mathrm{Uniform}(\mathrm{min}=0, \mathrm{max}=90),\\
  \sigma_\mathrm{gal} &\sim \mathrm{InverseGamma}(\alpha=2,\,\beta=20).
\end{align}

We also have the offset parameter $c$, and a measure of radial uncertainty $\sigma_r$:

\begin{align}
  c_\mathrm{arm} &\sim \mathrm{Cauchy}(\alpha=0,\,\beta=10),\\
  \sigma_r &\sim \mathrm{HalfCauchy}(\beta=0.2).
\end{align}

The Cauchy distirbution is equivalent to the Student's t-distribution with one degree of freedom, and was chosen due to its fatter tails than the normal distribution. Our likelihood function is therefore

\begin{equation}
\mathcal{L} \sim \mathrm{Normal}\left(\exp\left[\theta_\mathrm{arm}\tan\phi_\mathrm{arm} + c_\mathrm{arm}\right], \sigma_r\right).
\end{equation}

To perform inference, we make use of the No-U-Turn-Sampler (NUTS, \citealt{2011arXiv1111.4246H}), implemented in PYMC3\footnote{\url{https://docs.pymc.io/}}, an open source probabilistic programming framework written in python \citep{pymc3_paper}. We scale the radii of deprojected points to have unit variance, to aid the convergence of MC chains.

>>>>

Spiral arm points are deprojected to a face-on orientation using the disk inclination and position angle obtained through photometric model fitting performed in \Lingard. Arms are individually corrected to all have the same chirality (a pitch angle greater than or equal to zero) using the logarithmic spiral fit in \Lingard.

@@

We assume that a galaxy has a single value for pitch angle, $\phi_\mathrm{gal}$, and that the pitch angles of spiral arms in this galaxy, $\phi_\mathrm{arm}$, are constants (giving logarithmic spirals) drawn from a normal distribution centred on $\phi_\mathrm{gal}$, with some spread $\sigma_\mathrm{gal}$ common to all galaxies. We truncate this normal distribution between the physical limits of {0\degree} (a ring) and {90\degree} (a ``spoke''), giving

@@

Furthermore, we assume that the observed points in a \textit{Galaxy Builder} spiral arm, once deprojected, follow a logarithmic spiral with gaussian radial error $\sigma_r$,

\begin{equation}
\widetilde{r_\mathrm{arm}} = \exp\left(\widevec{\theta_\mathrm{arm}}\tan\phi_\mathrm{arm} + c_\mathrm{arm}\right).
\end{equation}

Where $\widetilde{r_\mathrm{arm}}$ is the model's predictions for the radii of the deprojected points in a \textit{Galaxy Builder} arm ($\widevec{r_\mathrm{arm}}$), and $\widevec{\theta_\mathrm{arm}}$ is the polar angles of the points.

We choose hyperpriors over $\phi_\mathrm{gal}$, $\sigma_\mathrm{gal}$, $c_\mathrm{arm}$ and $\sigma_\mathrm{r}$ of

\begin{align}
  \phi_\mathrm{gal} &\sim \mathrm{Uniform}(\mathrm{min}=0, \mathrm{max}=90),\\
  \sigma_\mathrm{gal} &\sim \mathrm{InverseGamma}(\alpha=2,\,\beta=20),\\
  c_\mathrm{arm} &\sim \mathrm{Cauchy}(\alpha=0,\,\beta=10),\\
  \sigma_r &\sim \mathrm{InverseGamma}(\alpha=2, \beta=0.5).
\end{align}

The inverse gamma distribution is used to aid the convergence of the Hamiltonian Monte Carlo algorithm used (discussed later). The Cauchy distirbution is equivalent to the Student's t-distribution with one degree of freedom, and was chosen due to its fatter tails than the normal distribution. Our likelihood function for $N$ arms, each of which with $n_\mathrm{arm}$ points is

\begin{equation}
  \mathcal{L} = \prod_{\mathrm{arm}=1}^{N}\left(2\pi\sigma_r^2\right)^{-n_\mathrm{arm}/2}
  \exp\left(\frac{||\widevec{r_\mathrm{arm}}\,-\,\widetilde{r_\mathrm{arm}}||^2}{2\sigma_r^2}\right).
\end{equation}

To perform inference, we make use of the No-U-Turn-Sampler (NUTS, \citealt{2011arXiv1111.4246H}), implemented in PYMC3\footnote{\url{https://docs.pymc.io/}}, an open source probabilistic programming framework written in python \citep{pymc3_paper}. To aid the convergence of MC chains, we scale the radii of deprojected points to have unit variance.

### 2020-04-14 sections/results.tex
separates analysis of galaxy pitch angle and arm pitch angle

% !TEX root = ../main.tex
\subsection{Constraints on Galaxy Pitch angle}
Our hierarchical model identifies the pitch angle of individual arms ($\phi_\mathrm{arm}$) with a high degree of certainty (uncertainty less than {1.6\degree} for 95\% of arms, assuming no error on disc inclination and position angle), however the pitch angle of a galaxy as a whole ($\phi_\mathrm{gal}$) is not well constrained. This is primarily a result of only having pitch angles measurements for a small number of arms per galaxy, and reflects the difficulty in providing a single value for the pitch angle of a galaxy containing individual arms with very different pitch angles. For galaxies with two arms identified in \textit{Galaxy Builder}, we have a mean uncertainty of {7.8\degree}, which decreases to {6.7\degree} and {5.8\degree} for galaxies with three and four arms respectively.

Our measure of inter-arm variability of pitch angle, $\sigma_\mathrm{gal}$ has a posterior distribution of $11.0\pm 1.0$, similar to the finding of \citet{2014ApJ...790...87D} and emphaising the need for fitting algorithms to not assume all arms have the same pitch angle.

\subsection{Spiral Winding}
\label{section:spiral_winding}
In order to test the possible progenitor distribution of our estimated galaxy pitch angles, we repeatedly perform an Anderson-Darling test over each draw present in the MCMC trace, resulting in a distribution of Anderson-Darling statistics. We will refer to this test as the \textit{marginalized Anderson-Darling test}. We also make use of the Kolmogorov-Smirnov test in a similar manner for comparison.

We perform the marginalized Anderson-Darling test for a potential source distribution uniform in $\cot\phi$ between the limits present in \citet{2019arXiv190910291P} ($1.00 < \cot\phi < 4.75$, or roughly $11.9^\circ < \phi < 45.0^\circ$). The resulting distribution of Anderson-Darling statistics can be seen in Figure \ref{fig:ad-cot-test}. We observe that we reject the null hypothesis at the 1\% level for 69\% of the possible realizations of galaxy pitch angle. While, with our sample and methodology, we cannot unilaterally reject winding of the kind described by \citet{2019arXiv190910291P}, it suggests this model is not found in our data.

\begin{figure*}
  \includegraphics[width=17.7cm]{plots/cot_uniform_marginalized_tests.pdf}
  \caption{The results of a marginalized Anderson-Darling test (top panel), with values corresponding to various confidence intervals shown, and marginalized Kolmogorov-Smirnov test p-values (lower panel). Moving rightwards on the x-axis implies greater confidence in rejecting the null hypothesis.}
  \label{fig:ad-cot-test}
\end{figure*}

Not seeing uniformity in $\cot\phi$ is not evidence against spiral winding, nor is it evidence for static spiral structures. Our results simply suggest that a more comprehensive test is needed for observational data of this kind.

\subsection{Dependence of pitch angle on Galaxy Morphology}
\label{section:morphology_comparision}
\subsubsection{Pitch angle vs. Bulge size}
We see no correlation between galaxy pitch angle derived from the hierarchical model and Galaxy Zoo 2's debiased \citep{Willett2013:1308.3496v2} \textit{pbulge}, which has been shown to be a good measure of bulge size.

We separate our sample into ``disc-dominated galaxies'' and ``obvious bulge galaxies'' using the debiased fractions from Galaxy Zoo 2 following \citet{2017MNRAS.469.3363K}, defining \textit{no bulge + just noticeable $>$ obvious + dominant} for the former and the converse for the latter. A marginalized two-sample Anderson-Darling test \citep{doi:10.1080/01621459.1987.10478517} does not find any evidence that the samples were drawn from different distributions; we do not reject the null hypothesis at the 1\% level for any samples. The distribution of Anderson-Darling test statistics is shown in the upper panel of Figure \ref{fig:ad-morphology-test}.

@@

>>>>

% !TEX root = ../main.tex
\subsection{Constraints on Galaxy Pitch angle}
Our hierarchical model identifies the pitch angle of individual arms ($\phi_\mathrm{arm}$) with less than {1.6\degree} of uncertainty for 95\% of arms, assuming no error on disc inclination and position angle. The pitch angle of a galaxy as a whole ($\phi_\mathrm{gal}$), however, is not well constrained. This is primarily a result of only having pitch angles measurements for a small number of arms per galaxy, and reflects the difficulty in providing a single value for the pitch angle of a galaxy containing individual arms with very different pitch angles. For galaxies with two arms identified in \textit{Galaxy Builder}, we have a mean uncertainty of ($\sigma_{\phi_\mathrm{gal}}$) of  {7.8\degree}, which decreases to {6.7\degree} and {5.8\degree} for galaxies with three and four arms respectively. This is consistent with the standard error on the mean for a galaxy with $N$ arms,

\begin{equation}
  \sigma_{\phi_\mathrm{gal}} = \frac{\sigma_\mathrm{gal}}{\sqrt{N}},
\end{equation}

where $\sigma_\mathrm{gal}$ is our measure of inter-arm variability of pitch angle and has a posterior distribution of $11.0^\circ\pm 1.0^\circ$. This variability is similar to the finding of \citet{2014ApJ...790...87D} and emphaises the need for fitting algorithms to not assume all arms have the same pitch angle.

\subsection{Dependence of pitch angle on Galaxy Morphology}
\label{section:morphology_comparision}
In order to test the possible progenitor distribution of our estimated arm pitch angles, we repeatedly perform an Anderson-Darling test \citep{10.2307/2286009} over each draw present in the MC trace, resulting in a distribution of Anderson-Darling statistics. We will refer to this test as the \textit{marginalized Anderson-Darling test}. We also make use of the two-sample Anderson-Darling \citep{doi:10.1080/01621459.1987.10478517} and Kolmogorov-Smirnov tests in a similar manner for comparison.

\subsubsection{Pitch angle vs. Bulge size}
We define a bulge prominence from Galaxy Zoo 2 as Equation 3 in \citet{2019MNRAS.487.1808M}:

\begin{equation}
  B_\mathrm{avg} = 0.2p_\mathrm{just noticeable} + 0.8p_\mathrm{obvious} + 1.0p_\mathrm{dominant}.
\end{equation}

We see no correlation between galaxy pitch angle derived from the hierarchical model and $B_\mathrm{avg}$. The Pearson correlation coefficient between the expectation value of galaxy pitch angle ($E(\phi_\mathrm{gal})$) and $B_\mathrm{avg}$ is -0.09. We see a slight correlation between $E(\phi_\mathrm{gal})$ and the fraction of volunteers who said a galaxy did not have a bulge (Pearson correlation coefficient of 0.19).

\comment{Do we need to do an Anderson-Darling test here?}

We separate our sample into ``weaker bulged galaxies'' and ``stronger bulged galaxies'' using , in order to examine whether they could be drawn from significantly different distributions. A marginalized two-sample Anderson-Darling test does not find any evidence that the samples were drawn from different distributions; we do not reject the null hypothesis at the 1\% level for any samples. The distribution of Anderson-Darling test statistics is shown in the upper panel of Figure \ref{fig:ad-morphology-test}.

@@

\subsection{Spiral Winding}
\label{section:spiral_winding}

For transient and reccurent spiral arms driven by self-gravity, \citet{2019arXiv190910291P} suggest that spiral patterns form at some maximum pitch angle ($\phi_\mathrm{max}$), continually wind up over time and finally dissapate at some minimum pitch angle ($\phi_\mathrm{min}$). They propose that, under a set of very simple assumptions, the evolution of pitch angle would be governed by

\begin{equation}
  \label{eq:winding}
  \cot{\phi} = \left[R\frac{\mathrm{d}\Omega_p}{\mathrm{d}R}\right](t - t_0) + \cot{\phi_\mathrm{max}},
\end{equation}

where $\Omega_p$ is the radially dependant pattern speed of the spiral arm and $t_0$ is the initial time at which it formed.

In QSDW theory, the pattern speed $\Omega_p$ is a constant in R, as spiral arms obey rigid-body rotation. If $\Omega_p$ instead varies with radius we would expect $\cot{\phi}$ to be uniformly distributed between $\cot{\phi_\mathrm{max}}$ and $\cot{\phi_\mathrm{min}}$.

In order to test this theory, \citet{2019arXiv190910291P} used a Kolmogorov-Smirnov test to examine the consistency of a sample of observed galaxy pitch angles with one uniform in cot. Pitch angles were measured using discrete fourier transformations in one- and two-dimensions, and as such do not account for inter-arm variations. They chose limits of $\cot{\phi} \in [1.00, 4.75]$ (roughly $11.9^\circ < \phi < 45.0^\circ$), motivated by examination of their data.

We aim to replicate their work here, using our sample and methods. We will make use of the marginalized tests described above, and examine winding on a per-arm basis, as well as a per-galaxy basis. Observation of the distribution of arm pitch angles in our sample suggests limits of $15^\circ < \phi < 50.0^\circ$.

\subsubsection{Galaxy Pitch angle}

Testing the uniformity in cot of $\phi_\mathrm{gal}$, between the limits specified above, using a marginalized Anderson-Darling test, results in rejecting the null hypothesis at the 1\% level for 68\% of samples. This suggests that a cot uniform model is not a good fit to galaxy pitch angle, but is not conclusive enough for us to unilaterally reject winding governed by Equation \ref{eq:winding}. The full distribution of Anderson-Darling statistics can be seen in the upper panel of Figure \ref{fig:ad-cot-test}.

This inconclusive result is perhaps unsurprising; were we to assume that spiral arms are transient and reccurent instabilities, there is little reason for all of the arms to be at precisely the same evolutionary stage at the same time. This is supported by the large spread in inter-arm pitch angles.

\subsubsection{Arm Pitch angle}
If we assume that spirals form and wind independantly in a galaxy and that their evolution over time can be described by Equation \ref{eq:winding}, the distribution of pitch angles of individual arms should be uniform in cot between our limits.

Using the marginalized Anderson-Darling test we cannot reject the null hypothesis at the 10\% level for any of the possible realizations of arm pitch angle. The resulting distribution of Anderson-Darling statistics in the lower panel of Figure \ref{fig:ad-cot-test}. This result is highly consistent with the model for spiral winding proposed by \citet{2019arXiv190910291P}, and can be seen as evidence that spirals are formed through local disc perturbations, and are primarily governed by local forces.

\begin{figure*}
  \includegraphics[width=17.7cm]{plots/combined_cot_uniform_marginalized_tests.pdf}
  \caption{The results of a marginalized Anderson-Darling test for uniformity in $\cot$ for $\phi_\mathrm{gal}$ (top panel) and $\phi_\mathrm{arm}$ (bottom panel), with values corresponding to various confidence intervals shown. Moving rightwards on the x-axis implies greater confidence in rejecting the null hypothesis.}
  \label{fig:ad-cot-test}
\end{figure*}

This result is highly sensitive to the lower limit of $\phi_\mathrm{gal}$ chosen; decreasing it to 10\degree results in us rejecting the cot-uniform progenitor at greater than the 0.1\% level for every realization of the posterior. As we have no information available on the biases present in \textit{Galaxy Builder} spiral arm classification, we should use the higher limits.

\comment{We do not account for observation effects, instead assuming that galaxy builder spiral arms are equally likely to be identified and recovered at all pitch angles. We also assume that the galaxy builder is representative of the general spiral population, which is not necessarily the case}
